Motional Quantum Ground-State Cooling Outside the Lamb-Dicke Regime

### 2017-06-25 paper.tex
References

Systems of individual neutral atoms trapped in optical tweezers are an exciting platform to study
quantum information, quantum chemistry, and quantum simulation of many-body systems \fxnote{REF}.
This approach provides full single-site resolution and control,
and can feature long-range interactions by using Rydberg atoms, polar molecules,
or dipolar atoms \fxnote{REF}.
Polar molecules in particular offer the additional advantages of long-lived internal states
and a high degree of tunability, including the shape, range,
and strength of the interaction \fxnote{REF}.
Optical tweezers can also be re-arranged in real-time,
allowing rapid preparation of atoms in complex geometries with high fidelity \fxnote{REF}.

In order to achieve long coherence times and full quantum control of the system,
it is typically necessary to cool the atom to the
three dimensional motional ground state in the optical tweezer.
This has been demonstrated for neutral Rb\cite{Thompson2013,Kaufman2012}
and Cs\cite{Li2012,Liu2017}
in experiments that had low initial temperature and small Lamb-Dicke parameter.
These features may not be easily achievable for other systems,
such as directly laser-cooled polar molecules \fxnote{REF},
or atoms with challenging sub-Doppler cooling such as Na, K, and Li \fxnote{REF}.
In this letter, we cool single sodium atoms trapped in optical tweezers to
the motional ground state with Raman sideband cooling.
Despite having a large Lamb-Dicke parameter and high initial temperature,
we are able to achieve a ground state probability of $77(4)\%$
by utilizing several new cooling techniques and a carefully optimized cooling sequence.
Our approach is quite general, and opens up ground-state cooling for other systems.

>>>>

Systems of individual neutral atoms trapped in optical tweezers are an exciting platform to study
quantum information, quantum chemistry, and quantum simulation of many-body systems\cite{Schlosser2001,DeMille2002,Muller2014Rydberg}.
This approach provides full single-site resolution and control,
and can feature long-range interactions by using Rydberg atoms, polar molecules,
or dipolar atoms\cite{Micheli2006,Muller2014Rydberg,Walker2012}.
Polar molecules in particular offer the additional advantages of long-lived internal states
and a high degree of tunability, including the shape, range,
and strength of the interaction\cite{Sortais2007,Beugnon2007,Ni2009}.
Optical tweezers can also be re-arranged in real-time,
allowing rapid preparation of atoms in complex geometries with high fidelity
\cite{Carpentier2013,Weiss2004}.

In order to achieve long coherence times and full quantum control of the system,
it is typically necessary to cool the atom to the
three dimensional motional ground state in the optical tweezer.
This has been demonstrated for neutral Rb\cite{Thompson2013,Kaufman2012}
and Cs\cite{Li2012,Liu2017}
in experiments that had low initial temperature and small Lamb-Dicke parameter.
These features may not be easily achievable for other systems,
such as directly laser-cooled polar molecules\cite{Shuman2010,Chae2017},
or atoms with challenging sub-Doppler cooling such as Na, K, and Li\cite{Shahriar1993a,Colzi2016,Salomon2013,Salomon2015}.
In this letter, we cool single sodium atoms trapped in optical tweezers to
the motional ground state with Raman sideband cooling.
Despite having a large Lamb-Dicke parameter and high initial temperature,
we are able to achieve a ground state probability of $77(4)\%$
by utilizing several new cooling techniques and a carefully optimized cooling sequence.
Our approach is quite general, and opens up ground-state cooling for other systems.
